---
title: "Boston Crime EDA"
author: "Pranali Bhosale"
date: "2/19/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Packages required:
readr, tidyverse, maps, mapproj, measurements, lubridate, tinytex, dplyr, ggplot2, tinytex 

```{r}
#library(maps)
library(lubridate)
library(readr)
library(tidyverse)
library(dplyr)
library(ggplot2)
#library(tinytex)
```




I have chosen the Boston Crime Incident dataset \n

TITLE: Crime Incident Reports (June 2015 - To Date) (Source: New System)\n

DATASET DESCRIPTION: Crime incident reports are provided by Boston Police Department (BPD) to document the initial details surrounding an incident to which BPD officers respond. This is a dataset containing records from the new crime incident report system, which includes a reduced set of fields focused on capturing the type of incident as well as when and where it occurred. Records in the new system begin in June of 2015.\n

```{r}
#Importing the raw data
crime_raw <- read.csv("C:/Users/prana/Desktop/SECOND YEAR/Data Viz/Lecture 4/crime.csv", na.strings="")

#Getting the summary of all variables in dataset
summary(crime_raw)

#Displaying the first 10 rows of the dataset
head(crime_raw, 10)
```



##Exploring the data

```{r}
#Checking for total number of missing values in variables
sum(is.na(crime_raw$OFFENSE_CODE))
sum(is.na(crime_raw$YEAR))
sum(is.na(crime_raw$MONTH))
sum(is.na(crime_raw$HOUR))
sum(is.na(crime_raw$REPORTING_AREA))
sum(is.na(crime_raw$Long))
sum(is.na(crime_raw$Lat))
sum(is.na(crime_raw$DISTRICT))
sum(is.na(crime_raw$SHOOTING))
sum(is.na(crime_raw$OFFENSE_CODE_GROUP))
sum(is.na(crime_raw$UCR_PART))
sum(is.na(crime_raw$STREET))

#Converting OCCURRED_ON_DATE variable to datetime format
crime_raw$OCCURRED_ON_DATE <- as_datetime(crime_raw$OCCURRED_ON_DATE)

#Replacing the District numbers with district names for convenience purpose
names_district = c(
A1 = 'Downtown',
A15= 'Charlestown',
A7= 'East Boston',
B2= 'Roxbury',
B3= 'Mattapan',
C6= 'South Boston',
C11= 'Dorchester',
D4= 'South End',
D14= 'Brighton',
E5= 'West Roxbury',
E13= 'Jamaica Plain',
E18= 'Hyde Park')
crime_raw$DISTRICT <- as.factor(names_district[(crime_raw$DISTRICT)] )

head(crime_raw,10)
```

We see that the missing values in variables in this dataset are as follows:\n
Numerical variables:\n
REPORTING_AREA : 30425
Long: 27709
Lat: 27709\n


Categorical variables:\n
SHOOTING: 425093
DISTRICT: 2306
OFFENSE_CODE_GROUP: 28080
UCR_PART: 28190
STREET: 15204\n



Let us first see the distribution for each variable in the crime dataset.I am using a function to plot histograms for contious variables and Bar plots for categorical vairables. To keep this report short, I am commenting the print command. (These graphs are present on my poster)\n


```{r}
#Writing a function called 'plotCols' which iterates through each column of a dataset and plots the distribution for each column.
plotCols <- function(anydataset) {
  #For loop is used to iterate through each column in the dataset.
  for(i in names(anydataset)) {
    #Checking to whether the column is continous or not.If found continous then histogram is plotted.
    if(is.numeric(anydataset[[i]])){
     g<- ggplot(data=anydataset, mapping= aes_string(x=i)) + 
       geom_histogram(color="black", fill="purple") 
     #print(g)
    }#if ends
    #If the Column is categorical then a bar plot is created as follows.
    else {
     #i <- factor(i)
     g<- ggplot(data=anydataset, mapping= aes_string(x=i)) + 
       geom_bar(color="black", fill="pink") 
     #print(g)
    }#else ends
  }#for loop ends
}#function ends

#Excluding three colums (Location, OCCURRED_ON_DATE, INCIDENT_NUMBER) before passing dataframe to plotting function.
graph<- crime_raw[, c(-17, -8, -1)]

#Calling the function to plot graphs for all columns in the dataset.
plotCols(graph)
```

Let's start looking at some charts now.                                                                                                           

```{r} 
#Boxplot for all variables
boxplot(drop_na(crime_raw) , main="Box plot for variables", col = "skyblue")

#Graph showing the number of incidents in Boston on a yearly basis
ggplot(crime_raw, aes(x = factor(YEAR))) + geom_bar(color="black", fill="purple") + 
ggtitle(label = "Total number of Offense in Boston by Year") +  xlab("Year")

#Crime incidents by Hour of the day using District as an added aesthetic
ggplot((crime_raw), mapping = aes(x=factor(HOUR), fill=DISTRICT)) + 
ggtitle(label = "Crime count by Hour") + geom_bar() + xlab("Hour")

#Crime incidents by Month of the day using District as an added aesthetic
ggplot((crime_raw), mapping = aes(x=factor(MONTH), fill=DISTRICT)) + 
ggtitle(label = "Crime count by Month") + geom_bar() + xlab("Month")

#Crime incidents by day of Week of the day using District as an added aesthetic
ggplot((crime_raw), mapping = aes(x=factor(DAY_OF_WEEK), fill=DISTRICT)) + 
ggtitle(label = "Crime count by Day of Week") + geom_bar() + xlab("Day of Week")

#Graph presenting the types of offense in Boston on a yearly basis
ggplot(data = (crime_raw), aes(x= factor(OFFENSE_CODE_GROUP), fill = factor(YEAR))) + geom_bar() + 
ggtitle(label = "Types of Offense in Boston by Year") + xlab("Offense Code Groups") + coord_flip()

#Graph showing the count of UCR Part crimes each year. UCR Part one crimes are the most dangerous kind including murder, rape etc. 
ggplot(data=drop_na(crime_raw), aes(x=factor(UCR_PART), fill=factor(UCR_PART), na.rm = TRUE)) + geom_bar() + 
  facet_wrap(~YEAR) + ggtitle("Count of UCR Part crimes each year") + 
  ylab("Count") + xlab("UCR Part crimes")  + coord_flip()

#Scatterplot using long and lat and using District as the added aesthetic
ggplot((crime_raw), aes(x=Lat, y=Long, fill=factor(DISTRICT), color=factor(DISTRICT))) + geom_point() + 
  scale_y_continuous(limits = c(-71.175,-71)) + scale_x_continuous(limits = c(42.225,42.400)) 

#Comparing days of the week with street
crime_raw %>% filter(fct_explicit_na(STREET) %in% (crime_raw  %>% count(STREET) %>% arrange(-n) %>% head(10) %>% pull(STREET))) %>% 
    ggplot(aes(DAY_OF_WEEK, fill = STREET)) + geom_bar(position = "fill") +
    coord_flip() + labs(x="Day of Week", title = "Comparing days of the week with street") + scale_fill_ordinal() 

#Comparing District with street
crime_raw %>% filter(fct_explicit_na(STREET) %in% (crime_raw  %>% count(STREET) %>% arrange(-n) %>% head(10) %>% pull(STREET))) %>% 
    ggplot(aes(DISTRICT, fill = STREET)) + geom_bar(position = "fill") +
    coord_flip() + labs(x="District", title = "Comparing District with street") + scale_fill_ordinal() 

#Graph showing the count crimes in District and filled with the offense description
drop_na(crime_raw) %>% filter(OFFENSE_DESCRIPTION %in% (crime_raw  %>% count(OFFENSE_DESCRIPTION) %>% arrange(desc(n)) %>% head(5) %>% pull(OFFENSE_DESCRIPTION))) %>% 
    ggplot(aes(DISTRICT, fill = OFFENSE_DESCRIPTION))+
    geom_bar()+
    coord_flip()+
    scale_fill_ordinal() +ggtitle("Count of Districts with offense description ")

#Graph showing the count crimes in District and filled with the Day of week aesthetic
drop_na(crime_raw) %>% filter(DAY_OF_WEEK %in% (crime_raw  %>% count(DAY_OF_WEEK) %>% arrange(desc(n)) %>% head(5) %>% 
                                                  pull(DAY_OF_WEEK))) %>% 
    ggplot(aes(DISTRICT, fill = DAY_OF_WEEK))+
    geom_bar()+
    coord_flip()+
    scale_fill_ordinal() +ggtitle("Count of Districts with Day of week ")

#Graph between two numerical variables - OFFENSE_CODE and REPORTING_AREA
ggplot(crime_raw) + geom_bin2d(aes(x=(OFFENSE_CODE), y=(REPORTING_AREA))) + ggtitle("Offense code vs Reporting area") + 
  labs(x="Offense code", y="Reporting area")

#Graph between two numerical variables - MONTH and HOUR
ggplot(crime_raw) + geom_bin2d(aes(x=factor(MONTH), y=factor(HOUR))) + ggtitle("MONTH vs HOUR") + labs(x="MONTH", y="HOUR")

```


```{r}
##TABLES
#Displaying the top 10 streets with most criminal incident reports
Top_streets<-(crime_raw %>% count((STREET)) %>% arrange(desc(n)))
print(head(Top_streets,10))

#Displaying the top 10 Districts with most criminal incident reports
Top_districts<- (crime_raw) %>% count((DISTRICT)) %>% arrange(desc(n))
print(head(Top_districts,10))

#Displaying the top 10 Offense code group with most criminal incident reports
Top_groups<- (crime_raw) %>% count((OFFENSE_CODE_GROUP)) %>% arrange(desc(n))
print(head(Top_groups,10))

#Displaying the top 10 Offense code group by Hour with most criminal incident reports
Top_gs<- (crime_raw) %>% group_by(fct_explicit_na(OFFENSE_CODE_GROUP), HOUR)%>% count() %>% arrange(desc(n))
print(head(Top_gs,10))

#Displaying the top 10 Offense code group by street with most criminal incident reports
Top_gd<- drop_na(crime_raw) %>% group_by(fct_explicit_na(OFFENSE_CODE_GROUP), STREET)%>% count() %>% arrange(desc(n))
print(head(Top_gd,10))

#Displaying the top 10 Offense descriptions by district with most criminal incident reports
Top_des_dist<- drop_na(crime_raw) %>% group_by(DISTRICT,fct_explicit_na(OFFENSE_DESCRIPTION))%>% count() %>% arrange(desc(n))
ggplot(Top_des_dist, aes(x=Top_des_dist$DISTRICT, fill=Top_des_dist$`fct_explicit_na(OFFENSE_DESCRIPTION)`)) + 
  geom_bar() + labs(title = "Count of crimes by district using offense description ", x="District", y="Offense Description")

#Displaying the lowest 10 Offense code group with most criminal incident reports
low_groups<- (crime_raw) %>% count((OFFENSE_CODE_GROUP)) %>% arrange((n))
print(head(low_groups,10))

#Displaying the top 10 Offense code descriptions with most criminal incident reports
Top_des<- (crime_raw) %>% count((OFFENSE_DESCRIPTION)) %>% arrange(desc(n))
print(head(Top_des,10))

#Displaying the top 10 REPORTING_AREA with most criminal incident reports
Top_area<- (crime_raw) %>% count((REPORTING_AREA)) %>% arrange(desc(n))
print(head(Top_area,10))
```

